---
title: "Hospitalization Data"
output: rmarkdown::github_document
---

```{r setup, include=FALSE}
# pull in prepared data
knitr::opts_chunk$set(echo = TRUE, warning = FALSE, message = FALSE)
library(dplyr)
library(ggplot2)

```

## CDC Hospitalization data

Hospitalization Rates and Characteristics of Patients Hospitalized with Laboratory-Confirmed Coronavirus Disease 2019 — COVID-NET, 14 States, March 1–30, 2020

Accessed from <https://www.cdc.gov/mmwr/volumes/69/wr/mm6915e3.htm?s_cid=mm6915e3_w> on April 17, 2020.

"Among 167 patients with available data, the median interval from symptom onset to admission was 7 days."

This paper is stored in the references directory as mm6915e3_H.pdf.

The paper references data that has been published on the following site. <https://gis.cdc.gov/grasp/covidnet/COVID19_3.html>

Currently, the script CDC_HospData.R compiles the downloaded data locally and stores a slim csv for plotting.
 
```{r CDCplot, echo=FALSE}
hospdata <- read.csv("../data/CDCHosp/hospdata.csv")

CDCplot <- hospdata %>%
  filter(CATCHMENT != "Entire Network") %>%
  ggplot(aes(MMWR.WEEK, WEEKLY.RATE, col = AGE.CATEGORY)) + 
  geom_line() + 
  facet_wrap(~ CATCHMENT) + 
  theme_minimal() +
  ylab("Hospitalization per 100k residents") +
  ggtitle("Cumulative hospitalizations")
#  xlab("")
#  scale_color_paletteer_d("awtools::a_palette")

CDCplot
```

## IHME Hospitalization predictions

IHME predictions were pulled on three separate dates in April: April 4, 9, and 12. 

Currently, the script IHME_HospPreds.R compiles the data locally and stores a slim csv for plotting only the states available in the CDC data.

```{r IHMEplot, echo=FALSE}
IHMEplotdata <- read.csv("../data/IHMEProjections/IHMEplotdata.csv")

CDCStates <- c("Utah", "Tennessee", "Oregon", "Ohio", "New York", "New Mexico", "Minnesota", "Michigan", "Maryland", "Iowa", "Georgia", "Connecticut", "Colorado", "California")

IHMEplotdata <- IHMEplotdata %>%
  filter(location_name %in% CDCStates,
         doy < 120,
         doy >= 70)

IHMEtime <- ggplot(IHMEplotdata, aes(doy, HospRate, col = location_name)) + 
  geom_line() + 
  facet_wrap(~ dataseries) + 
  theme_minimal() +
  ylab("Hospitalization per 100k residents") 

IHMEstate <- ggplot(IHMEplotdata, aes(doy, HospRate, col = factor(dataseries))) + 
  geom_line() + 
  facet_wrap(~ location_name) + 
  theme_minimal() +
  ylab("Hospitalization per 100k residents") +
  ggtitle("Current beds required")

IHMEstate
```

